Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems

### 2015-09-08 Docs/Paper/LaTex/template.tex
major improvements for abstract and introduction

\begin{abstract}
We present an alternative approach to solve the hardware (HW) and software (SW) partitioning problem, which uses Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open MultiProcessing. The multi-core SMT-based BMC approach allows initializing many verification instances based on the number of available processing cores. Each instance checks for a different optimum value until the optimization problem is satisfied. The goal is to show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. 
We implement and evaluate our algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). We also compare our approaches to other state-of-the-art optimization tools: Matlab and Z3. Experimental results show that no single optimization tool solves all benchmarks; however, Matlab and Z3 are the most efficient ones to solve HW-SW partitioning problems, although ESBMC had a significant performance improvement from its sequential to parallel version.
\end{abstract}

@@

Nowadays, with the strong development of embedded systems, the design phase plays an important role. At early stages, the design is split into separated flows: hardware and software. Consequently, the partitioning decision process, which deals with the decisions upon which parts of the application have to be designed in hardware (HW) and which in software (SW), must be supported by any well-structured methodology. If not, this leads to a number of issues (design flow interruptions, redesigns, and undesired iterations) which affects the overall development process, the quality and the lifecycle of the final system. Starting at the 1990s, intensive research was performed, and several approaches proposed, as shown in ~\cite{Arato2003} and ~\cite{Mann2007}.

In any HW and SW design of complex systems, more time is spent on verification than on construction ~\cite{Baier2008}. Formal methods based on model checking offer great potential to obtain a more effective and faster verification in the design process. Programs may be viewed as mathematical objects with behavior that is, in principle, well determined. This makes it possible to specify programs using mathematical logic, which constitutes the intended (correct) behavior. Then, one can try to give a formal proof or otherwise establish that the program meets its specification ~\cite{Clarke2009}. Research in formal methods has led to the development of very promising verification techniques, which facilitate the early detection of errors. Model-based verification techniques use models that describe the possible system behavior in a mathematically precise and unambiguous manner. The system models are accompanied by algorithms that systematically explore all the states of the system model.

In ~\cite{Trindade2015} and ~\cite{Trindade2014} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) to perform HW-SW partitioning in embedded systems. The present work extends those studies since there is a substantial improvement in terms of the genetic algorithm and the SMT-based verification method, which has been extended with a multi-core architecture. Multi-core processors have been used in all segments of industry to implement high-performance computing ~\cite{Wu2014}. In particular, hardware platforms, together with multi-processing platforms, have allowed verification algorithms to distribute tasks executions across multiple processors, which generate an increase in performance if compared to single-core solution. However, most verification algorithms still disregard the limitations of the CMOS technology, which limits the increase of the chip’s frequency after it reaches 4 GHz.

Here, we exploit the availability of multi-core processors; in particular, SMT-based verification methods are applied to the HW-SW partition problem and experimental results are compared to ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and also the $\nu$Z tool, which is a state-of-art optimization tool based on SMT~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the Optimization Toolbox of Matlab~\cite{OpenMP1998}, while $\nu$Z is a built-in tool to the SMT solver Z3 in which its main function is to check the satisfiability of logical formulas based on objective functions. To the best of our knowledge, this is the first work to use a multi-core SMT-based verification to solve a HW-SW partitioning problem in embedded systems. We implement our ideas with the Efficient SMT-based Bounded Model Checker (ESBMC) tool ~\cite{Cordeiro2012}. As its main contribution, this paper shows that it is possible to take advantage of an SMT-based BMC tool in a multi-core architecture to solve optimization problems.

@@

%5
\bibitem{Trindade2015}
Trindade, A., Cordeiro. L. (2015), ``{Applying SMT-based verification to hardware/software partitioning in embedded systems. In: Design Automation for Embedded Systems (to appear).}''

@@

>>>>

\begin{abstract}
We present an alternative approach to solve the hardware (HW) and software (SW) partitioning problem, which uses Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open Multi-Processing. In a nutshell, the multi-core SMT-based BMC approach allows initializing many verification instances based on the number of available processing cores. Each instance checks for a different optimum value until the optimization problem is satisfied. We implement and evaluate our algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). We also integrate the $\nu$Z tool into ESBMC, which is a state-of-art optimization tool based on SMT. We compare all our approaches to another state-of-the-art optimization tool (Matlab). Experimental results show that no single optimization tool solves all HW-SW partitioning benchmarks; however, Matlab and $\nu$Z are the most efficient ones to solve HW-SW partitioning problems, although ESBMC had a significant performance improvement in particular cases.
\end{abstract}

@@

Nowadays, with the strong development of embedded systems, the design phase plays an important role. At early stages, the design is split into separated flows: hardware and software. The partitioning decision process, which deals with decisions upon which parts of the application have to be designed in hardware (HW) and which in software (SW), must be supported by any well-structured methodology. If not, this leads to a number of issues (design flow interruptions, redesigns, and undesired iterations) which affects the overall development process, the quality, and the lifecycle of the final system. Starting at the 1990s, intensive research was performed, and several approaches proposed, as shown in ~\cite{Arato2003} and ~\cite{Mann2007}.

In any HW and SW design of complex systems, more time is spent on verification than on construction ~\cite{Baier2008}. Formal methods based on model checking offer great potential to obtain a more effective and faster verification in the design process. Software (and systems) may be viewed as mathematical objects with behavior that is, in principle, well determined. This makes it possible to specify software using mathematical logic, which constitutes the intended (correct) behavior. Then, one can try to give a formal proof or otherwise establish that the software meets its specification ~\cite{Clarke2009}. Research in formal methods has led to the development of very promising verification techniques, which facilitate the early detection of errors. Model-based verification techniques use models that describe the possible system behavior in a mathematically precise and unambiguous manner. The system models are accompanied by algorithms that systematically explore all the states of the system model.

In~\cite{Trindade2015,Trindade2016,Trindade2014} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) to perform HW-SW partitioning in embedded systems. The present work extends those studies since there is a substantial improvement in terms of the SMT-based verification method, which has been extended with a multi-core binary-search approach, as well as, the integration of the $\nu$Z tool into ESBMC, which is a state-of-art optimization tool based on SMT~\cite{Bjorner2015}. Multi-core processors have been used in all segments of industry to implement high-performance computing~\cite{Wu2014}. In particular, hardware platforms, together with multi-processing platforms, have allowed verification algorithms to distribute tasks executions across multiple processors, which generate an increase in performance if compared to single-core solution~\cite{Holzmann11}. However, most algorithms for software verification still disregard the limitations of the CMOS technology and focus their solution on a single-core approach~\cite{Cordeiro2012,Clarke2004,Armando2009,Ganai2006}.

Here, we exploit the availability of multi-core processors; in particular, SMT-based verification methods are applied to the HW-SW partition problem in three different ways using multi-core ESBMC with OpenMP:  ESBMC-MC using a sequential-search, ESBMC-PS using a parallel-search, and ESBMC-PB using a binary-search. Experimental results are compared to ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and also the $\nu$Z tool, which supports only a single-core approach~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the Optimization Toolbox of Matlab~\cite{OpenMP1998}, while $\nu$Z is a built-in tool to the SMT solver Z3 in which its main function is to check the satisfiability of logical formulas based on objective functions. All multi-core ESBMC with OpenMP approaches, together with $\nu$Z, are implemented with the Efficient SMT-based Bounded Model Checker (ESBMC) tool~\cite{Cordeiro2012}.

\textbf{Contributions.} The main contribution of the present study is to show that it is possible to take advantage of an SMT-based BMC tool in a multi-core architecture to solve optimization problems. Experimental results show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. Additionally, we integrate the MaxSMT solver $\nu$Z into an off-the-shelf BMC tool that is typically used for software verification. To the best of our knowledge, this is the first work to use a multi-core SMT-based verification and a MaxSMT solver to check for HW-SW partitioning problems in embedded systems.

\textbf{Availability of Data and Tools. }Our experiments are based on a set of publicly available benchmarks. All benchmarks, tools, and results of our evaluation are available on a supplementary web page\footnote{http://esbmc.org/}.

@@

%5
\bibitem{Trindade2015}
Trindade, A., Cordeiro. L. (2015), ``{Applying SMT-based verification to hardware/software partitioning in embedded systems,}'' In: Design Automation for Embedded Systems (to appear).

%5
\bibitem{Trindade2016}
Trindade, A. B., Ismail, H. I., Cordeiro, L. C. (2015), ``{Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems (short paper),}'' To appear in V Brazilian Symposium on Computing Systems Engineering (SBESC), pp. 1-4.

@@

%44
\bibitem{Holzmann11}
Gerard J. Holzmann, Rajeev Joshi, Alex Groce (2011), {``Swarm Verification Techniques''}, In: IEEE Trans. Software Eng. v. 37, n. 6, pp. 845--857.
